---
title: "Intro R Answer Key - Week 2"
author: "Dr. Dani Kirsch & Dr. Jaina Agan"
date: "`r Sys.Date()`"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)
```

```{r packages, include = FALSE}
library(ggplot2)
```

## Dataset Description

For this week's optional practice problems, we have provided another dataset for you to use.  

`midwest` is a dataset that’s available in R as part of the `ggplot2` pacakge, similar to the `complete_old` dataset we have been using from the `ratdat` package. The `midwest` dataset has demographic information for counties from several Midwest states from the 2000 U.S. Census.

1. Write and execute the following code to read about where the dataset came from and what the variables are.  

```{r data-descr, message = FALSE}
?midwest
```

2. Use `str` to learn more about data types in the `midwest` dataset. Of the primary vector types we discussed in this week's workshop (character, integer, numeric, logical), which are represented in this dataset?  

```{r str-midwest, echo=TRUE, include = TRUE, message = FALSE}
str(midwest)
```

## Make a Fancy Boxplot

Create a graph from the `midwest` data that compares population density between states. Follow the instructions below:  

1. Compare `state` on the x-axis to `popdensity` on the y-axis.  
2. Create a combined boxplot and scatter plot graph with the following features:  
    * Use `geom_jitter` to create the scatter plot layer.  
      * Set the color of the scatter plot data points to vary by the variable `state`.  
      * Choose a new point shape (`shape = `). Shape is identified using integers, and these are some of your point shape options:    
![point shape options in R][pch_img]
      * Set a transparency level.  
      * Make the point size larger (`size = `).  
    * Use `geom_boxplot` to create the boxplot layer.  
      * Remove the `fill` color.  
      * Remove the `outlier.shape` to avoid double-plotting outliers.  
3. Set a new theme (e.g., `theme_classic()`, `theme_bw()`)  
4. Change the labels so that:  
    * Plot title is "Midwest Population Demographics (2000)".  
    * X-axis is "State".  
    * Y-axis is "Population Density (person/unit area)".    
5. Change other plot features using the `theme` function so that:  
    * The size of the plot title text is 16 and the text face is **bold**.  
    * The position of the legend is "none" (removes legend).  

The graph you create should look similar to the following graph:  

```{r fancy-boxplot, echo=TRUE, warning = FALSE}
ggplot(data = midwest,
       aes(x = state,
           y = popdensity)) +
  geom_jitter(aes(color = state),
              shape = 18,
              alpha = 0.6,
              size = 3) +
  geom_boxplot(outlier.shape = NA,
               fill = NA) +
  theme_bw() +
  labs(title = "Midwest Population Demographics (2000)",
       x = "State",
       y = "Population Density (person/unit area)") +
  theme(plot.title = element_text(size = 16,
                             face = "bold"),
        legend.position = "none")
  
```

## Summary Statistics

Compute the following calculations on the `midwest` dataset:  

1. What is the minimum value of population density (`popdensity`)?  
```{r min-pop-den, echo=TRUE}
min(midwest$popdensity)
```

2. What is the maximum value of total population (`poptotal`)?  
**Hint**: Use the function `max()`  
```{r max-pop, echo=TRUE}
max(midwest$poptotal)
```

3. What are the quantiles (25%, 50%, 75%) for the percent of a county's population that is college educated (`percollege`)?     
```{r quant-college, echo=TRUE}
quantile(midwest$percollege, prob = c(0.25,0.5,0.75))
```

4. What is the average number of adults per county (`popadults`)?  
**Hint**: Use the function `mean()`  
```{r avg-adult, echo=TRUE}
mean(midwest$popadults)
```

## Sequences

1. Create a sequence that runs from 8 to 85 at intervals of 7. Store this sequence as an object called `weird_seq`.  
```{r weird-seq, echo=TRUE}
weird_seq <- seq(from = 8, to = 85, by = 7)

# print out list values
weird_seq
```
2. Create a sequence that runs from 1900 to 2025 at intervals of 5. Store this sequence as an object called `year_seq`.  
```{r year-seq, echo=TRUE}
year_seq <- seq(from = 1900, to = 2025, by = 5)

# print out list values
year_seq
```
3. Create a sequence that is runs from 12 to 22 and has a length of 47 (47 total items in the sequence). Store this sequence as an object called `seq_length`.  
```{r seq-rep, echo=TRUE}
seq_length <- seq(from = 12, to = 22, length.out = 47)

# print out list values
seq_length
```

## Reflections

1. What was the most challenging aspect of this week’s workshop? Were you able to overcome it? If not, what assistance do you need to continue working through it?  

2. What was the most rewarding aspect of this week's workshop?  

[pch_img]: pch.png "Point Shapes"